---
title: "Class 12"
author: "Gary Le"
date: "May 11, 2018"
output: github_document
---


# Hands-on

* Loading in Bio3D and HIV-1 Protease
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

* Reading PDB file into R
```{r}
hiv <- read.pdb(file.name)
```

* Trimming large PDB object into separate components for easier access
```{r}
#Trimming files by protein/ligand
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

#Writing out trimmed files
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, "1hsg_ligand.pdb")
```

* Preparing PDB files in AutoDocTools.
    + AutoDocTools used to add Hydrogen and charge data back to the file since the visualization technique for the protein didn't have enough resolution to resolve hydrogens.
    + But, we already know where hydrogens _should_ be based off of biochemical knowledge.
    
* AutoDocTools Grid > GridBox settings
    + Spacing (angstrom) = 1.000
    + (x,y,z) Box size = (30, 30, 30)
    + (x,y,z) center = (16, 25, 4)

* Scored docking with 1hsg_ligand using AutoDoc Vina
    + config = config.txt
    + log = log.txt
    + Results stored in all.pdbqt
    
* Using Bio3D to prepare a PDB file for VMD to read
```{r}
library(bio3d) 
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

