---
title: "Class 17"
author: "Gary Le"
date:: "May 30th, 2018"
output: github_document
---

* Loading Cytoscape 3 and testing connection.
```{r}
library(RCy3)
library(igraph)
library(RColorBrewer)

cytoscapePing()
cytoscapeVersionInfo()
```

* Importing sample data for network
```{r}
#Loading data from link
prok_vir_cor <- read.delim("https://bioboot.github.io/bggn213_S18/class-material/virus_prok_cor_abundant.tsv", stringsAsFactors = FALSE)

#Verifying head
head(prok_vir_cor)
```

* Set data frame for graph
```{r}
g <- graph.data.frame(prok_vir_cor, directed = FALSE)

# N.B. igraph needs to be loaded
class(g)

g

# N.B. "UNW- 845 1544" means that the graph has 854 nodes, 1544 edges and is an Undirected, Named, and Weighted graph 
```

```{r}
plot(g)

# "This is a hot-mess! Lets turn of the blue text labels"
# - Dr. Grant
```

```{r}
plot(g, vertex.label=NA)

# Still cheese balls.
```

```{r}
plot(g, vertex.size=3, vertex.label=NA)
```

* Optionally, graph with ggplot
```{r}
library('ggraph')

ggraph(g, layout = 'auto') +
  geom_edge_link(alpha = 0.25) +
  geom_node_point(color="lightblue")
```

* Accessing vertex and edge data with V() and E() respectively
```{r}
V(g)
E(g)
```

* Detecting communities by measuring betweenness
```{r}
cb <- cluster_edge_betweenness(g) #Default is Girvan-Newman method
cb
```

* Mapping communities onto network
```{r}
plot(cb, y=g, vertex.label=NA,  vertex.size=3)

# Original plot for comparison
plot(g, vertex.size=3, vertex.label=NA)

```

* membership() can be used to see which community each node is apart of
```{r}
head( membership(cb) )
```

* Other useful functions are to see the degrees in the network
```{r}
# Calculate and plot node degree of our network
d <- degree(g)
hist(d, breaks=30, col="lightblue", main ="Node Degree Distribution")
```
* As we can see, the histogram looks like a a scale-free network.

```{r}
plot( degree_distribution(g), type="h" )
```

*Now, we can take a look at centrality. This helps gauge the importantness of a node. Even Google uses it! We'll be using the Google PageRank method
```{r}
pr <- page_rank(g)
head(pr$vector)
```

* Visualizing graph based on importance of nodes by centrality
```{r}
# Make a size vector btwn 2 and 20 for node plotting size
library('BBmisc')
v.size <- BBmisc::normalize(pr$vector, range=c(2,20), method="range")

# Plotting with vertex size
plot(g, vertex.size=v.size, vertex.label=NA)
```

* We can also visualize centrality with the easier metric of degrees each node has.
```{r}
# Setting vertex size on a distribution of [2,20] based off degrees
v.size <- BBmisc::normalize(d, range=c(2,20), method="range")

plot(g, vertex.size=v.size, vertex.label=NA)
# Notice how some nodes change in importance/size while others don't
```

* A little more complex method is to use betweenness which, roughly, divides the shortest paths through a node by its edges.
```{r}
# Another useful and aptly named function.
b <- betweenness(g)
v.size <- BBmisc::normalize(b, range=c(2,20), method="range")

plot(g, vertex.size=v.size, vertex.label=NA)
# Note the drastic increase in size of the path linking the two neighborhoods on the right of the graph
```

### Annotating Network with Taxonomical Information
```{r}
# Loading in phage id data
phage_id_affiliation <- read.delim("https://bioboot.github.io/bggn213_S18/class-material/phage_ids_with_affiliation.tsv")
head(phage_id_affiliation)
```

```{r}
# Loading in bacterial id data from SILVA DB
bac_id_affi <- read.delim("https://bioboot.github.io/bggn213_S18/class-material/prok_tax_from_silva.tsv", stringsAsFactors = FALSE)
head(bac_id_affi)
```

* Preparingnetwork for Cytoscape Visualization
```{r}
# Extract out our vertex names
## The gene names are stored as the vertex attributes (this set only has names)
genenet.nodes <- as.data.frame(vertex.attributes(g), stringsAsFactors=FALSE)

head(genenet.nodes)
```

```{r}
# Using grep to see how many phage nodes we have (Nodes that start with "ph_")
length( grep("^ph_",genenet.nodes[,1]) )

## 764 Phage nodes
## 81 non-phage nodes
```

```{r}
# Merging different columns to make a classification table

# We dont need all annotation data so lets make a reduced table 'z' for merging
z <- bac_id_affi[,c("Accession_ID", "Kingdom", "Phylum", "Class")]

# Our final, merged table with 
n <- merge(genenet.nodes, z, by.x="name", by.y="Accession_ID", all.x=TRUE)
head(n)
```
    + All the Bacteria should be properly labeled now.

    + Moving onto the more complicated phage IDs
```{r}
# Check on the column names before deciding what to merge i.e. what we have/need to use
colnames(n)
```

```{r}
# We only want the phages that match
colnames(phage_id_affiliation)
```

```{r}
# Again we only need a subset of `phage_id_affiliation` for our purposes
y <- phage_id_affiliation[, c("first_sheet.Phage_id_network", "phage_affiliation","Tax_order", "Tax_subfamily")]

# Add the little phage annotation that we have
x <- merge(x=n, y=y, by.x="name", by.y="first_sheet.Phage_id_network", all.x=TRUE)

## Remove duplicates from multiple matches
x <- x[!duplicated( (x$name) ),]
head(x)
```
```{r}
# Storing our node data
genenet.nodes <- x

```

* Final preparations for Cytoscape
```{r}
# Adding edge data from our graph "g""
genenet.edges <- data.frame(igraph::as_edgelist(g))
names(genenet.edges) <- c("source",
                          "target")

# Adding the edge weights which are just another edge attribute
genenet.edges$Weight <- igraph::edge_attr(g)[[1]]

# Adding the proper names; strictly specifiying type as character
genenet.edges$source <- as.character(genenet.edges$source)
genenet.edges$target <- as.character(genenet.edges$target)
genenet.nodes$id <- as.character(genenet.nodes$name)

# Final plot for Cytoscape. Takes a simple list of nodes and edges
## N.B. RCy3 has been updated and cyPlot() is now a defunct fxn
## Use this method instead to create Cytoscape Windos (cw)
### The method takes data frames as inputs
cw <- createNetworkFromDataFrames(nodes = as.data.frame(genenet.nodes), 
                                  edges = as.data.frame(genenet.edges), 
                                  title = "Tara Oceans")
```

* Coloring the graph 
```{r}
# N.B. English use of 'colour' vs 'color' b/c Dr. Grant is Irish
families_to_colour <- unique(genenet.nodes$Phylum)
families_to_colour <- families_to_colour[!is.na(families_to_colour)]

node.colour <- RColorBrewer::brewer.pal(length(families_to_colour), "Set3")

# Setting more color rules
setNodeColorMapping("Phylum",
                 families_to_colour,
                 node.colour,
                 "lookup",
                 default.color = "#ffffff")
```

* Saving Image
```{r}
# exportImage(filename="net2",
#           type="png",
#           h=350)

#Access Denied error in commandsPOST
```



