---
title: "Class 13"
author: "Gary Le"
date: "May 16th, 2018"
output: github_document
---

## Loading in 1000 Genomes Project Data

* Reading in data from Mexican in Los Angeles population (MXL) of 1000 Genome Projects 
```{r}
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

* Sample should have 64 rows (patients)

```{r}
#Table out different genotypes
table(genotype[,2])

#Percent Frequency of Genotypes
table(genotype[,2]) / nrow(genotype) * 100
```

* Loading Libraries to read sequencing scores from fastqsanger files
```{r}
library(seqinr)
library(gtools)

#test values
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred 
```

## Analyzing Population Scale genomics data

* Reading in and summarizing population data for SNP
```{r}
pop <- read.table("https://bioboot.github.io/bggn213_S18/class-material/
rs8067378_ENSG00000172057.6.txt")

summary(pop)

```

```{r}
boxplot(pop)
```

* Checking expression value of samples with G|G homozygous and other genotypes
    + Shows that A|A genotype has the highest expression followed by heterozygous
```{r}
summary(pop$exp[ pop$geno == "A/A"])
summary(pop$exp[ pop$geno == "A/G"])
summary(pop$exp[ pop$geno == "G/G"])

```

```{r}
#Box plots of expression by genotype (exp ~ geno) from the population data.
pop.box <- boxplot(exp ~ geno, data = pop, notch = TRUE)
```

* Fancy version of boxplot by ggplot2
```{r}
#install.packages("ggplot2")
# N.B. ggplot only takes data frames
library("ggplot2")

#Customize graph aesthetic by modifying the aes() arguments.
# N.B. aes() takes the input in the form of (factor,data)
# N.B. the + geom_foo() tells ggplot which type of geometry our plot should have
ggplot(pop, aes(geno,exp)) + 
  geom_boxplot(notch = TRUE, outlier.shape = NA)
```

```{r}
sessionInfo()
```

